---
title: "Automatic vs. Manual Transmission - Regression Modelling"
author: "Robert Jonczy"
date: "23.08.2015"
output: pdf_document
---

# Introduction

In this research performed on `mtcars` dataset I will try to answer following questions:

- “Is an automatic or manual transmission better for MPG” 
- “Quantify the MPG difference between automatic and manual transmissions”

# Data Transformations

First I load data and perform some data transformations to present some featuers as factors rather than numeric:

```{r}
data(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am, labels = c("Automatic", "Manual"))
```

# Data Exploratory
```{r}
str(mtcars[,c('mpg', 'am')])
head(mtcars[,c('mpg', 'am')])
cor(datasets::mtcars)["mpg",]
aggregate(mpg ~ am, data = mtcars, mean)
```
As shown from data manual tranmissions appear to achieve a higher MPG rating as opposed to those of automatic transmissions.

Letting the null hypothesis be: automatic transmissions have a better MPG rating vs manual transmissions. We shall determine if this true by calculating the P-value.

```{r}
t.test(mtcars$mpg ~ mtcars$am, conf.level=0.95)
```

The p-value 0.001374 is small (less than 0.05) and confidence interval does not contain 0 shows we have to **reject null hypothesis**. There is significant difference Automatic and Manual transmission, we can say that cars with automatic transmission have lower mpg than manual transmission.


# Regression Modelling

First we will try to fit a simple model with `mpg` as outcome and `am` as predictor.

```{r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)
```

Based on the Model only **33.8%** of the variance is explained using the coefficient of determination.

As a next step lets try to fit a model to all variables, and look at the p-values.

```{r}
fit.step <- step( lm(mpg ~ . , data = mtcars), direction = "both", trace = 0, steps=100 )
summary(fit.step)
```

Based of the model 84 percent of variation of mpg vs tramission type is explained via the coefficient of determination, R2.

R^2 is 0.84 means 84% of the variability explained by this model. Next, we compare linear model mpg ~ am with the best model using ANOVA a.


```{r}
bestfit <- lm(mpg ~ am + wt + qsec, data = mtcars)
anova(fit, bestfit)
summary(bestfit)
```

The model explains 84% of difference between mpg and tranmission type; furthermore with an exceptionally low P-value one can conclude that the null hypothesis:Automatic transmissions have a better MPG rating vs Manual transmissions can be rejected, i.e. automatic tranmissions do not achieve a better MPG rating. 


# Conclusion

The manual transmision is better than automatic by 1.81. MPG will decrease by 2.49 per 1000 lb. MPG will decrease with bigger number of cylinders

# Appendix

## Plot1: mpg histogram and kernel density
```{r, echo=FALSE, fig.align='center'}
par(mfrow = c(2,1))
hist(mtcars$mpg, breaks = 10, xlab = "MPG", main = "MPG histogram")
plot(density(mtcars$mpg), main = "kernel density", xlab = "MPG")
```

## Plot2: mpg by transmission type
```{r, echo=FALSE, fig.align='center'}
par(mfrow = c(1,1))
boxplot(mpg ~ am, data = mtcars,
        col = c("blue", "green"),
        xlab = "Transmission Type",
        ylab = "Miles per Gallon",
        main = "MPG by Transmission Type")
```

## Plot3: residuals
```{r, echo=FALSE}
par(mfrow = c(2,2))
plot(bestfit)
```